---
title:"AI"
author: "Dongdong Kong"
date: "`r Sys.Date()`"
output: html_document
---

# Referencen

1. Lian, X., Piao, S., Chen, A., Huntingford, C., Fu, B., Li, L. Z. X., et al. (2021). Multifaceted characteristics of dryland aridity changes in a warming world. Nature Reviews Earth & Environment, 2(4), 232–250. <https://doi.org/10.1038/s43017-021-00144-0>


```{r}
source("scripts/main_pkgs.R")
library(sf)
library(sf2)
library(lattice.layers)

sp_cont <- list("sp.lines", get_continent(), lwd = 0.2, first = F)
sp_layout <- list(sp_cont)
# sp_sign = list("sp.polygons", poly_shade, first = FALSE, lwd = 0.1)
load(file = "data-raw/Multimodel mean E Trends from CIMP6 and GLEAM.RData")
AI <- raster("INPUTS/AridityIndex_MSWEP-prcp_div_GLEAM-Ep_1980-2020.tif") %>%
  raster2array() %>%
  as.numeric()
```

# Figure S1:

准备数据：
```{r}
library(rgdal)
library(lattice.layers)
library(rcolors)
library(sf)

sp_cont <- list("sp.lines", get_continent(), lwd = 0.5, first = F)
sp_layout <- list(sp_cont)

g <- readGDAL("INPUTS/AridityIndex_MSWEP-prcp_div_GLEAM-Ep_1980-2020.tif")
g$band1 %<>% clamp(c(0, 10), TRUE) # %>% cut(brks)
```

## Dryland的划分标准

"Drylands are traditionally defined by the United Nations Environment Programme 
as areas with AI ≤ 0.65, and can be further subcategorized into dry sub-humid 
(0.65 > AI ≥ 0.5), semi-arid (0.5 > AI ≥ 0.2), arid (0.2 > AI ≥ 0.05) and 
hyper-arid (AI < 0.05) regions." (**Lian 2021, Box1**)

$$AI = Prcp / PET$$

Dryland: AI <= 0.65

| Category     |  Definition |
| ---- | ---- |
|dry sub-humid | 0.65 > AI ≥ 0.5 |
|semi-arid     | 0.5 > AI ≥ 0.2  |
|arid          | 0.2 > AI ≥ 0.05 |
|hyper-arid    | AI < 0.05       |

出图：
```{r}
# brks = c(-Inf, 0, 0.05, 0.2, 0.5, 0.65, Inf)
# brks2 = c(0.05, 0.2, 0.5, 0.65, 1, 2, 5)
brks <- c(-Inf, 0.2, 0.5, 0.65, Inf)
levs <- c("arid", "semi-arid", "sub-humid", "humid")

brks2 <- c(0.05, 0.2, 0.65, 1)
brks <- c(-Inf, 0.05, 0.2, 0.5, 0.65, 1:5, 20, Inf)
brks <- c(-Inf, 0.05, 0.2, 0.5, 0.65, Inf)

nbrk <- length(brks) - 1
n_more <- 2
cols <- get_color(rcolors$amwg256, nbrk + n_more) %>%
  .[-(1:n_more)] %>%
  rev()
p <- sp_plot(g,
  brks = brks,
  cols = cols,
  xlim = c(-180, 240),
  ylim = c(-60, 90),
  aspect = 0.5,
  sp.layout = sp_layout,
  key.num2factor = TRUE
) +
  layer_title(x = 0, y = 1, labels = c("AI = P/PET")) +
  layer_latFreq(
    bbox = c(185, 240, -60, 90), zlim = c(-2, 2),
    col.regions = NULL,
    unit = "native", is_spatial = TRUE
  ) +
  layer_barchart(x = 0.01, y = 0.05, width = 0.22) +
  layer_contourf(at = brks2, col = "black")
# layer_statistic(x = 0.4, y = 0.1, cex = 1.4)
write_fig(p, "Figure_S01_AridityIndex_SpatialDistribution2.pdf", 10.7, 5)
```
